\section{Initiating problem}
When it comes to game-development the work can be divided into levels.
Low-level is the basic workings of a game; physics, graphics and so on.
High-level is for example the behaviour of entities within the game world.
One of the most popular language to construct a game-engine with is C++, which is 'multi-paradigm' and a 'general purpose'-kind, meaning that it is created with general software development in mind.\\
For high-level development, the popular languages are of the scripting-kind like \emph{Lua}, which is a 'multi-paradigm' language.
For the languages that are 'general-purpose', frameworks are created to extend the functionality of each language, so that the various aspects of a video game are handled correctly and efficiently.
The popular scripting languages like Lua and JavaScript are as well 'general-purpose' and therefore no exception.
This means that countless hours of work are spent to extend functionality of a language, not tailored for this specific task, and the construction of game-engines is needed to save developers time on the low-level implementation.\\
The problem here is that there are few to none programming languages created solely for video game development, and as the before-mentioned tools and frameworks indicate, these kinds of languages would be beneficial to game-developers.
Another problem is the fact that most games are programmed imperatively, making development harder for beginners and non-programmers.
Since role-playing games are increasingly popular throughout the world, an RPG-focused programming language could be a good place to start.
Which is our initiating problem.\\
\emph{"There is currently no language specialised in describing turn-based battles in computer role-playing games}
\begin{comment}
\emph{"While the game-development industry is continuously growing and role-playing games are so popular, why isn't there an RPG-based programming language available?"}
\end{comment}

%TODO Needs refurbishing, comments for fixes below....

%
%Why is it necessary to program imperative to develop games?
%Tala um low-level (c++ game engine stuff)
%Tala um high-level (scripting.. .sem er -það sem við erum að gera)
%Reyna að koma aðeins meira inn á það svæði og benda á að tungumál eins og LUA og JavaScript eru 'general purpose' alveg eins og hin og þar af leiðandi gæti það verið nice að ný focus tungumál í staðinn fyrir framework.....




%In this project, we want to design and construct our own programming language along with the tools to translate and make use of the designed language.
%This presents us with following tasks:
%\begin{itemize}
%	\item Designing a language
%	\item Designing a fitting grammar for the language
%	\item Designing and implementing an engine to showcase the capabilites of the language
%	\item Designing and implementing a compiler to translate the language for our engine
%\end{itemize}

%This brings us to an initial problem statement:\\
%\emph{\textbf{How do we efficiently design and implement a readible and writable language, capable of simplifying entity creation for a Role-playing game.}}